### 2016-06-28 paper.tex
More stuff

%"runningheads" enables:
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
%This is good for other readers to enable proper archiving among other papers and pointing to content.

%Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
\documentclass[runningheads,a4paper]{llncs}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

%better font, similar to the default springer font
%cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
\usepackage[%
rm={oldstyle=false,proportional=true},%
sf={oldstyle=false,proportional=true},%
tt={oldstyle=false,proportional=true,variable=true},%
qt=false%
]{cfr-lm}
%
%if more space is needed, exchange cfr-lm by mathptmx
%\usepackage{mathptmx}

\usepackage{graphicx}

%extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

%put figures inside a text
%\usepackage{picins}
%use
%\piccaptioninside
%\piccaption{...}
%\parpic[r]{\includegraphics ...}
%Text...

%Sorts the citations in the brackets
%\usepackage{cite}

\usepackage[T1]{fontenc}

%for demonstration purposes only
\usepackage[math]{blindtext}

%for easy quotations: \enquote{text}
\usepackage{csquotes}

%enable margin kerning
\usepackage{microtype}

%tweak \url{...}
\usepackage{url}
%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
\makeatletter
\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
\makeatother
\urlstyle{same}
%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother

%diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
%slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
%\usepackage{diagbox}

%required for pdfcomment later
\usepackage{xcolor}

% new packages BEFORE hyperref
% See also http://tex.stackexchange.com/questions/1863/which-packages-should-be-loaded-after-hyperref-instead-of-before

%enable hyperref without colors and without bookmarks
\usepackage[
%pdfauthor={},
%pdfsubject={},
%pdftitle={},
%pdfkeywords={},
bookmarks=false,
breaklinks=true,
colorlinks=true,
linkcolor=black,
citecolor=black,
urlcolor=black,
%pdfstartpage=19,
pdfpagelayout=SinglePage,
pdfstartview=Fit
]{hyperref}
%enables correct jumping to figures when referencing
\usepackage[all]{hypcap}

%enable nice comments
\usepackage{pdfcomment}
\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}

%compatibality with TODO package
\newcommand{\todo}[1]{\commentatside{#1}}

%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}

\usepackage{xspace}
%\newcommand{\eg}{e.\,g.\xspace}
%\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.,\ }
\newcommand{\ie}{i.\,e.,\ }

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
    <-6>  MnSymbolC5
   <6-7>  MnSymbolC6
   <7-8>  MnSymbolC7
   <8-9>  MnSymbolC8
   <9-10> MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage[normalem]{ulem}

% \def\OP#1{{\color{red}OP: \it #1}}
% \def\ODdel#1{\bgroup\markoverwith{\textcolor{red}{\rule[0.5ex]{2pt}{1pt}}}\ULon{#1}}
\def\OP#1{#1}  % do nothing - disable the comments in whole documment

\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\def\OD#1{{\color{darkgreen}OD: \it #1}}
\def\ODdel#1{\bgroup\markoverwith{\textcolor{darkgreen}{\rule[0.5ex]{2pt}{1pt}}}\ULon{#1}}
%\def\OD#1{\relax} % make my comments vanish
%\def\ODdel#1{#1} % make my comments vanish

\begin{document}

%Works on MiKTeX only
%hint by http://goemonx.blogspot.de/2012/01/pdflatex-ligaturen-und-copynpaste.html
%also http://tex.stackexchange.com/questions/4397/make-ligatures-in-linux-libertine-copyable-and-searchable
%This allows a copy'n'paste of the text from the paper
\input glyphtounicode.tex
\pdfgentounicode=1

\title{A Dataset of Operator-client Dialogues Aligned with Database Queries}

\author{Ondřej Plátek \and Filip Jurčíček}
\institute{Charles University in Prague, Faculty of Mathematics and Physics \\
Institute of Formal and Applied Linguistics \\
Malostranské náměstí 25, 11800 Praha 1, Czech Republic \\
\email{\{oplatek,jurcicek\}@ufal.mff.cuni.cz}}

\iftrue % FIXME uncomment for final submission
\author{Firstname Lastname \and Firstname Lastname }
\institute{
Insitute\\
\email{anonym@anonymous}
}
\fi

\maketitle
% \OP{Fill missing values about the dataset collected so far} \\
% \OP{grammar check} \\
\OD{Přemýšlím, jestli ten název neudělat nějak míň popisný a víc vysvětlující (něco o end-to-end dialogue systems)?}

\begin{abstract}
    This paper presents a novel dataset for training end-to-end task oriented conversational agents.
    The dataset contains conversations between an operator – a task expert, and a client who seeks information about the task.
    Along with the conversation transcriptions, we record database API calls performed by the operator, which capture a distilled meaning of the user query.
    We expect that the easy-to-get supervision of database calls will allow us to train end-to-end dialogue agents with significantly less training data.
 	The dataset is collected using crowdsourcing and the conversations cover the well-known restaurant domain.
    Quality of the data is enforced by mutual control among contributors.  % TODO and manually verified on random sub-sample 
    The dataset is available for download under the Creative Commons 4.0 BY-SA license.
\end{abstract}

\vspace{-1.00em}
\keywords{task-oriented dialogue, end-to-end, dataset}
\vspace{-1.00em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
\vspace{-0.50em}
We present a new dataset of human-human task-oriented conversations in the restaurant information domain, suitable for supervised training of autonomous systems. 
Currently, the dataset contains only \OP{62} dialogues because we are still collecting more data, but we estimate we will have more than 700 dialogues until the camera ready deadline.
\OD{We are currently collecting data for the set using crowdsourcing. We have collected 62 dialogues so far; more than 700 dialogues are planned for the final version of the dataset.\footnote{\OD{We expect to release the set before the camera-ready deadline.}}}

We aim at representing the conversation in a way that makes it easy to train an automated system to replace a human operator. 
In contrast to previously released datasets, our dataset contains easy-to-collect high-quality transcriptions of the client and operator actions during conversations.\OD{tady úplně nevím, k čemu se kontrastuješ}
We introduce a~very little overhead by collecting database calls in addition to the transcription of the conversation.
In fact, we mimic very closely a~real situation in call centers where operators search for answers through a database user interface. 
Just logging the calls to the task database provide us with information of a similar quality as a full manual dialogue state annotation, but not annotated incrementally after each turn.\OD{od “but” nerozumím}
Tracking the database calls in task oriented systems together with conversation transcription should not introduce significant overhead during the data collection, but should presumably boost end-to-end conversational agent as much as annotated dialogue state when training with such data.\OD{but should provide a similar amount of supervision for training an end-to-end conversational agent as using fully annotated dialogue state ??}

Current Spoken Dialogue Systems (SDS) are either handcrafted and use no training data~\cite{duvsek2014alex,raux2005let} but require non-trivial amount of expert work, or they are gradually improved from initial policy through live user interaction~\cite{young2010hidden,gasic2011line}. Recently, SDS have also been trained using supervised learning in an end-to-end manner~\cite{wen2016network,williams2016end}.\footnote{The~work of \cite{williams2016end} also fine-tuned the conversation with reinforcement learning after the supervised learning stage.}
To train a system using reinforcement learning, one only needs to collect enough live user conversation containing explicit feedback.
However, the feedback signal is harder to interpret than supervised annotation: Thousands of live conversations are required to improve a rather simple policy.\cite{gasic2011line}
Supervised learning methods reduced the number of conversations needed for training a reasonable policy down to a few hundreds\cite{wen2016network}, but require explicit annotation for all components used in the dialogue system.
To our knowledge, all current systems require full dialogue state annotation as a minimum~\cite{wen2016network,young2010hidden}.\OD{Tuhle větu asi do footnotu?}
With the dialogue state being a simplification of the dialogue history with no broadly-accepted form, it is very expensive to collect such annotated data because the dialogue state typically differs among domains and its structure is hard to explain to annotators.
In this paper, we focus on collecting a supervised training set with annotation rich enough so that supervised models similar to~\cite{wen2016network} can be trained with a few hundreds of dialogues, yet much easier to collect than previous methods.

We believe that our dataset dataset can be used for first experiments in training end-to-end systems using supervised learning without explicit dialogue state annotation.
The restaurant information domain used in our dataset is well established thanks to dialogue state tracking (DST) challenges~\cite{williams2013dstc1,henderson2014dstc2,henderson2014dstc3}. 
The DST challenges proved that the domain is of a reasonable complexity.
In addition, the challenges showed that the DST task is solvable by architectures which may be extended to end-to-end conversational agents~\cite{platek2016recurrent,vodolan2015hybrid,henderson2014word}.
Note that \ODdel{we collected human-to-human dialogues on Cambridge restaurant domains, but} our data-collection method is \OD{completely} domain independent.

The paper is structured as follows: Section~\ref{sec:repre} motivates the annotation and data formats\OD{?} used in our dataset.
In Section~\ref{sec:collection}, we introduce our collection process using crowdsourcing.
Section~\ref{sec:props} presents the dataset properties, and we list related work in Section~\ref{sec:related}.
We conclude the paper and suggest future work in~Section~\ref{sec:conc}.

% \section{End-to-end training of dialogue systems}
% \label{sec:training}

\section{Supervised Feedback for Task-oriented Dialogue} \label{sec:repre}
\vspace{-0.50em}
We present a dataset for training a task-oriented conversational agent in a well-defined domain.\OD{In this paper, we assume that we need to train a...}
We assume that the agent has a task database\footnote{In the Cambridge restaurant domain the database is a simple table containing information about restaurants.} at hand and plays the role of the {\it operator}.
The other interlocutor in the conversation is a~human {\it client}.\OD{that seeks information stored in the database}
Clients seek information stored in the database and use the system, which replaced a~human operator, to obtain desired information.\OD{A tuhle větu už pak můžeš vyhodit.}

We focus on data collection for text-to-text task-oriented dialogue systems.
Traditionally, one would train the pipeline of components of task-oriented dialogue systems, i.e., automatic speech recognition (ASR), language understanding (LU) unit, dialogue tracker (DT), dialogue manager (DM), natural language generator (NLG) for text-to-text conversational agents and text-to-speech (TTS) module~\cite{duvsek2014alex}.
However, only the text-to-text pipeline of LU, DT, DM, and NLG is typically trained using just in-domain data~\cite{gasic2011line,jurvcivcek2014factored,duvsek2016sequence}.
The ASR~\cite{platek2014free} and the TTS~\cite{taylor2009text} modules are typically trained using larger datasets.
Our datasets contains data for training an~end-to-end system solving task of LU, DT, DM and NLG pipeline of traditional architecture.\OD{Tenhle odstavec bych ještě trochu překopal.}

% Me to nesedi jako related work - related work je o sbirani dat
% Tohle me prijde zduvodneni, ze takovy (maly, supervised) dataset dava smysl
% Zatim nechavam zde: \OD{následující dva odstavce zní dost jako “related work”... možná bych to tam částečně popřesouval? a tady nechal jen ten statement, že stačí málo trénovacích dat?}
It was shown that a few hundreds of annotated domain specific conversations are needed to train a relatively well performing LU~\cite{jurvcivcek2014factored} as well as Dialogue State Tracking~\cite{young2010hidden} components.\OD{“few” = málo, “a few” = pár, několik}
The work of~\cite{duvsek2016sequence} and~\cite{mairesse2010phrase} showed that only several hundreds of turns with appropriate annotation is needed for training a usable NLG unit in simple domains.
Even the crucial task of action selection can be trained using a POMDP dialogue manager\OD{přepsal bych na “using reinforcement learning”, protože nikde nedefinuješ, co je POMDP} and few thousands conversations despite the fact that rewards are only provided at the end of a dialogue~\cite{gasic2011line}.

We believe that end-to-end systems can also be trained using a few hundreds of conversations for narrow goal-oriented domains, such as Cambridge restaurant domain~\cite{henderson2014dstc2} or its simplification~\cite{wen2016network}, if additional supervision in the form of simple annotation is used.
The~work of~\cite{wen2016network} showed that only 680 conversations annotated just at the dialogue state level are needed for training a complete end-to-end system performing the task of summarizing the dialogue history, action selection, and response generation.

We realize that a more fine-grained annotation helps to train the models faster, but simply collecting more data is often more beneficial.\OD{citace? a nevím, jestli tady ta věta je nutná. Možná bych vynechal všechno až do “we believe that annotating...”??}
As a consequence, we regard the works~\cite{wen2016network,bordes2016learning,williams2016end} as an important step forward to reducing the need for rich annotation in dialogue agent training.
We believe that annotating the dialogue state according to an expert-handcrafted ontology is both an~artificial and a~very labor-intensive data collection task. Therefore, our work aims at collecting conversation with lighter and easier to get annotation.
We suggest recording conversation between a human operator and a human clients via a crowdsourcing platform, i.e., collecting cheap high quality training data.
A similar recording setup is known from call centers where the operator is provided with a database interface so that they quickly find relevant information and play the role of a domain expert.
For our dataset, we collect the operator's database calls instead of dialogue state annotation after each turn.\OD{Asi bych tohle dal hned po “eaiser to get annotation”. Pak tu větu “a similar recording setup...” a až na konec “we suggest crowdsourcing”, který bych změnil na “To obtain the dataset, we record conversations between human operators and human clients...”}

The calls to the database represent very accurately the client's intention, which is the original purpose of dialogue state annotation.
On the other hand, the database calls are not present in every turn and thus by logging only the calls, one loses the possibility to track partially expressed client's intention.
However, we argue that in high quality human-human conversation, there are few turns which are related to the domain and do not contain calls to the database, as demonstrated in~Section~\ref{sec:props}.
It is worth noting that logging the operator's database calls in operator-client dialogues is already a well-established practice in call centers, a~convenient field for commercial application of dialogue systems.

\section{Dataset Collection Process} \label{sec:collection}
\vspace{-0.50em}
Our data collection process aims at obtaining natural in-domain conversations.
We decided to collect the dataset within the~well studied Cambridge restaurant domain since previous datasets~\cite{williams2013dstc1,henderson2014dstc2,henderson2014dstc3} in that domain demonstrated that a meaningful and reasonable-length conversations can be held here.\OD{tady už se částečně opakuješ, na jednom místě to zkrať}
Also, it was proven that the domain is on one hand difficult enough that even the state-of-the art systems do not reach human-level performance in the role of the operator~\cite{henderson2014dstc2} and on the other hand, numerous systems~\cite{young2010hidden,gasic2011line,wen2016network} demonstrated their ability to provide information to the clients with a reasonable quality.

We define our domain by choosing the database from the 2nd Dialogue State Tracking Challenge (DSTC2)~\cite{henderson2014dstc2}.
We use the database for steering the dialogue of the two human interlocutors a~{\it client} and an~{\it operator}.\OD{tuhle větu bych vynechal}
We use the CrowdFlower (CF) crowdsourcing platform to collect the conversations.

We first describe the database which defines our domain in~Section~\ref{sec:db}.
Then we describe how the hired workers use our interface to play either the client or the operator role in~Sections~\ref{sec:client} and~\ref{sec:operator}, respectively.
Finally, in~Section~\ref{sec:async}, we describe how we collect conversations one response at a time without connecting the interlocutors in real-time.

\vspace{-1.00em}
\subsection{Domain database}
\label{sec:db}

We use exactly the same database as provided for the DSTC2 challenge~\cite{henderson2014dstc2}.
The database contains information about 107 restaurants in Cambridge, and their properties are stored in several columns.
We include in brackets the number of different values for each column:
\vspace{-0.50em}\OD{Když dáš dvousloupcovou tabulku, tak i ušetříš trochu místa}
\begin{itemize}  % tabulku misto tohohle?
    \item name (107),
    \item price range (4),
    \item area (7),
    \item food (25),
    \item address (101),
    \item phone number (97),
    \item postcode (62).
\end{itemize}
\vspace{-0.50em}

The address and the phone number are unique for each restaurant but are sometimes missing; the restaurant name is always known. 
There are typically multiple restaurants with the same postcode, price range, area, or food type properties; these properties might also be unknown for a restaurant.

We use this database to generate\OD{prepare} client goals for CF users (see Figure~\ref{fig:client}).
The database is also available through a simple interface to CF users playing the role of the operator (see Figure~\ref{fig:operator}).\OD{nechceš tu odkazovat na ty podsekce spíš než na ty obrázky?}

\vspace{-1.00em}
\subsection{Client interface}
\label{sec:client}

The client is asked to request a specific information – a goal from the operator.
He may choose not to seek the goal immediately but submit more appropriate response such as greeting instead.
The goals are displayed to the CF worker in the client role gradually turn-by-turn and the goals may contradict each other.
For example, the client may at first want a Chinese food, but later it changes her mind and wants to find restaurant serving British food.

Before submitting a reply the client is instructed to read the dialogue history and marked the goals which were presented to the operator as demonstrated in Figure~\ref{fig:client}.
By marking which goals have been already presented to the operator we force the CF workers to pay attention to the dialogue history before submitting their reply.
In Section~\ref{sec:async} we describe how we combine the workers replies to form a~conversations.

\begin{figure}
\vspace{-1.00em}
\begin{center}
\includegraphics[height=30em]{gui-annotators-client}
\caption{Client annotation interface}
\end{center}
\vspace{-1.00em}
\label{fig:client}
\end{figure}
\vspace{-1.00em}

\vspace{-1.00em}
\subsection{Operator interface}
\label{sec:operator}
The operator is asked to politely reply the clients needs according the information stored in the database.
In order to provide the information to clients the operator uses the DB interface depicted in~Figure~\ref{fig:operator}.
Her or she filters the restaurants using a full text search on the database.
The content of the database is intentional hidden if no filter is specified so the filter has to be used.
The CF workers are also instructed explicitly mark the rows of the restaurants about which provide the information in their answer.
Similarly to the client role, an operator does not have to reply to all requested information at once.
It is up to the~worker which information is provided first.

The full text search, our interface to the database, is implemented in Javascript as a simple substring matching over values in the database.
Multiple constrains can be expressed by two substrings separated by a comma.
We used such a simple interface because it is easy to generate and also reconstruct the candidate results given the query and the database.
The CF workers were instructed to mark a candidate row from the results about which they inform in their reply.

\begin{figure}
\vspace{-1.00em}
\begin{center}
\includegraphics[height=35em]{gui-annotators-system}
\caption{Operator annotation interface}
\end{center}
\vspace{-1.00em}
\label{fig:operator}
\end{figure}
\vspace{-1.00em}

\vspace{-1.00em}
\subsection{Asynchronous collection without real-time responses}
\label{sec:async}
We design the crowdsourcing jobs so the CF contributors are able to work independently to each other.
The roles of client and operator take turns as sets of partial conversations are submitted for collecting a single new response.
As a consequence, a~single conversation is collected from multiple contributors over several job submissions.

Each utterance in conversation may be submitted by different contributor, so contributors need to read carefully the dialogue history to understand the conversation before submitting their response.
The workers are able to mark dialogues which does not make sense and thus they self-assess quality of the conversations.
The contributors mark the conversations as finished if they have no more goals to fulfill and all the goodbyes were said.

We bootstrap the conversation either with an operators introduction "{\it Hello , welcome to the Cambridge restaurant system...}" or we bootstrap the conversation with a~full turn where we add clients greeting e.g. {\it "Hi"} to aforementioned operators utterance.
If we initialize the dialogue with a~full turn the first contributor plays role of the system if we bootstrap the conversation only with~the operators greeting, the contributor plays the client's role.

In a~single job submission, we let the workers append to each dialogue a~single response.
Note also that the conversations are typically rather short so full dialogues are collected after few rounds of submissions.
The workers are paid for every submitted response the same amount independently if they play client or operator role and they typically contribute to multiple conversation during one job submission.
The contributors are assigned roles randomly.

To avoid poor language level of English we restricted the CF worker to English-speaking countries.
We further perform heuristically check for too trivial answers and we added ability to the CF workers to rate each others submission.
It help us filter low quality conversations, but more importantly it motivated workers not to cheat.

\section{Dataset Properties} \label{sec:props}
\vspace{-0.50em}
We just started to collect our dataset and so far we collected only \OP{62} conversations.
The average length of a conversation is \OP{3.8} turns, meaning that the contributors in both roles reply in average circa three times before the conversation is finished\footnote{First system reply is automatically generated and for some conversation even the first client reply is also initialized as described in Section~\ref{sec:async}}.
During an average conversation \OP{2.4} tasks are requested and \OP{1.9} answered.
% Typically in \OP{65} \% case the conversation is ended by the client.
% As we mention the conversation are short, taking only from \OP{four to six turns} including first turn with greetings.
% \begin{table}
% \begin{center}
% \begin{tabular}{lrr}
% \hline
% Set   & \# Dialogues & \# Turns \\
% \hline
% % Joint  &  todo &  todo \\
% % /a/SSD/oplatek/e2end/log/2016-06-08-17-31-38.400-dstc-INDEP_labels-d0.7-w100-e100/
% % /a/SSD/oplatek/e2end/log/TEST-ORIGINAL-2016-06-08-17-31-38.400dstc-INDEP_labels-d0.7-w100-e100-reward-0.9080-step-0036224
% Train  &   ? & ? \\
% Dev &   ? & ? \\
% Test &   ? & ? \\
% \hline
% \end{tabular}
% \caption{Accuracy on development and test set}
% \vspace{-2em}
% \end{center}
% \label{tab:props}
% \end{table}
During the data collection process we discarded over \OP{35} \% of replies because the utterances were of low quality.
Note that we discard only the low quality utterance, the prefix of the conversation is submitted for finishing the conversation in next rounds of jobs.
Once we will collect enough conversation we intend to split them into training, development and test set. % (see Table~\ref{tab:props}).

\section{Related Work} \label{sec:related}
\vspace{-0.50em}
Our work is closely related  to work of~\cite{williams2013dstc1,henderson2014dstc2,henderson2014dstc3} in the sense that the conversation are held in the same domain.
On the other hand, the collection process differs substantially because we do not use any artificial system in the operator's role.  
We rather focus on collecting high quality human conversations. 
In addition, in our work we do not collect complicated annotations on LU and DST level.

A~relevant work of~\cite{wen2016network} collected human-human dialogues in similar but more restricted domain than the Cambridge domain.
Their dataset is not yet published so we compare only with their statistics.
The dataset is of similar size as we intend to collect and a~very similar collection scheme was used where Amazon Mechanical Turk workers submitted one reply at a time without being connected to each other in real-time.
On the other hand, they collected explicit dialogue state annotation which takes significant annotation effort.

Another line of research~\cite{vodolan2016data} used CF to collect human-human conversation for interactive-learning dataset.
However, the collected dialogues were later annotated by expert annotators which goes directly against our intentions to avoid any expensive annotations.

There is a~significant amount of work which used Wizard of Oz experiments~\cite{whittaker2002fish,walker1997evaluating,rieser2008learning} for studying linguistic properties of dialogues, evaluating proof of concept or even collecting bootstrap training data, but according our knowledge no other work than~\cite{wen2016network} have used cheap crowdsourcing workers to collect a~full training dataset for supervised learning of dialogue management or end-to-end conversations.

\section{Conclusion and Future work} \label{sec:conc}
\vspace{-0.50em}
We present a new human-human dialogue dataset with annotation of user intention in form of operator's database calls.
We also introduced a novel data collection setting which resembles work of operators in call centers and introduce minimum overhead to crowdsourcing workers.
% TO mi prijde ze v call -centrech delaji stejne\OD{, apart from stating their replies in natural language (?)}
The collected dialogues are published under Creative Commons 4.0 BY-SA license on available to download at {\it anonymized}.  % lindat\footnote{todo where}.

We plan to increase the number of collected dialogues till the camera ready deadline to 700 or more.
We intend to use the dataset to train and evaluate an~end-to-end conversational model which uses only the database calls as additional supervision to the recorded responses.

\vspace{-1.40em}
\subsubsection*{Acknowledgments}
We would like to thank Ondřej Dušek for useful comments.
This research was partly funded by the Ministry of Education, Youth and Sports of the Czech Republic under the grant agreement LK11221, core research funding, grant GAUK 1915/2015, and also partially supported by SVV project number 260 333. 
% We gratefully acknowledge the support of NVIDIA Corporation with the donation of the Tesla K40c GPU used for this research.
% Cloud computational resources were provided by the MetaCentrum under the program LM2010005 and the CERIT-SC under the program Center CERIT Scientific Cloud, part of the Operational Program Research and Development for Innovations, Reg.\ no. CZ.1.05/3.2.00/08.0144.

% ``something in quotes'' using plain tex or use \enquote{the enquote command}.
% cref Demonstration: Cref at beginning of sentence, cref in all other cases.
%
% \Cref{fig:simple} shows a simple fact, although \cref{fig:simple} could also show something else.
%
% \Cref{tab:simple} shows a simple fact, although \cref{tab:simple} could also show something else.
%
% \Cref{sec:intro} shows a simple fact, although \cref{sec:intro} could also show something else.
%
% Brackets work as designed:
% <test>
%
% The symbol for powerset is now correct: $\powerset$ and not a Weierstrass p ($\wp$).
%
% \begin{inparaenum}
% \item All these items...
% \item ...appear in one line
% \item This is enabled by the paralist package.
% \end{inparaenum}
% In the bibliography, use \texttt{\textbackslash textsuperscript} for ``st'', ``nd'', ...:
% E.g., \enquote{The 2\textsuperscript{nd} conference on examples}.
% When you use \href{http://www.jabref.org}{JabRef}, you can use the clean up command to achieve that.

% Winery~\cite{Winery} is graphical \commentontext{modeling}{modeling with one \enquote{l}, because of AE} tool.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-1.00em}
\bibliographystyle{splncs03}
\bibliography{paper}

% All links were last followed on June 28, 2016.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}